Tensor denoising and completion based on ordinal observations

### 2020-06-09 paper/ICML/supplements/latexsource/ordinalTv4.tex
remove space adjustment

\vspace{-.1cm}
\section{Model formulation and motivation}
\vspace{-.1cm}
\subsection{Observation model}
\vspace{-.1cm}
Let $\tY$ denote an order-$K$ $(d_1,\ldots,d_K)$-dimensional data tensor. Suppose the entries of $\tY$ are ordinal-valued, and the observation space consists of $L$ ordered levels, denoted by $[L]:=\{1,\ldots,L\}$. We propose a cumulative link model for the ordinal tensor $\tY=\entry{y_\omega}\in[L]^{d_1\times \cdots\times d_K}$. Specifically, assume the entries $y_\omega$ are (conditionally) independently distributed with cumulative probabilities:
\begin{equation}\label{eq:model}
\mathbb{P}(y_\omega\leq \ell)=f(b_\ell-\theta_\omega),\ \text{for all}\ \ell\in[L-1],
\end{equation}
where $\mb=(b_1,\ldots,b_{L-1})$ is a set of unknown scalars satisfying $b_1<\cdots <b_{L-1}$, $\Theta=\entry{\theta_\omega}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is a continuous-valued parameter tensor satisfying certain low-dimensional structure (to be specified later), and $f(\cdot):\mathbb{R}\mapsto[0,1]$ is a known, strictly increasing function. We refer to $\mb$ as the cut-off points and $f$ the link function.

@@

\vspace{-.1cm}
\subsection{Latent-variable interpretation}\label{sec:latent}
\vspace{-.1cm}

@@

\subsection{Problem 1: Tensor denoising}~\label{sec:denoising}
\vspace{-.1cm}
The first question we aim to address is tensor denoising:

@@

\section{Conclusions}
We have developed a low-rank tensor estimation method based on possibly incomplete, ordinal-valued observations. A sharp error bound is established, and we demonstrate the outperformance of our approach compared to other methods. The work unlocks several directions of future research. One interesting question would be the inference problem, i.e.,\ to assess the uncertainty of the obtained estimates and the imputation. Other directions include the trade-off between (non)convex optimization and statistical/computational efficiency. While we have provided numerical evidence for the success of nonconvex approach, the full landscape of the optimization remains open. The interplay between computational efficiency and statistical accuracy in general tensor problems warrants future research.

>>>>

\section{Model formulation and motivation}
\subsection{Observation model}
\vspace{-.1cm}
Let $\tY$ denote an order-$K$ $(d_1,\ldots,d_K)$-dimensional data tensor. Suppose the entries of $\tY$ are ordinal-valued, and the observation space consists of $L$ ordered levels, denoted by $[L]:=\{1,\ldots,L\}$. We propose a cumulative link model for the ordinal tensor $\tY=\entry{y_\omega}\in[L]^{d_1\times \cdots\times d_K}$. Specifically, assume the entries $y_\omega$ are (conditionally) independently distributed with cumulative probabilities:
\begin{equation}\label{eq:model}
\mathbb{P}(y_\omega\leq \ell)=f(b_\ell-\theta_\omega),\ \text{for all}\ \ell\in[L-1],
\end{equation}
where $\mb=(b_1,\ldots,b_{L-1})$ is a set of unknown scalars satisfying $b_1<\cdots <b_{L-1}$, $\Theta=\entry{\theta_\omega}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is a continuous-valued parameter tensor satisfying certain low-dimensional structure (to be specified later), and $f(\cdot)\colon \mathbb{R}\mapsto[0,1]$ is a known, strictly increasing function. We refer to $\mb$ as the cut-off points and $f$ the link function.

@@

\subsection{Latent-variable interpretation}\label{sec:latent}

@@

\subsection{Problem 1: Tensor denoising}~\label{sec:denoising}
The first question we aim to address is tensor denoising:

@@

\section{Conclusions}
We have developed a low-rank tensor estimation method based on possibly incomplete, ordinal-valued observations. A sharp error bound is established, and we demonstrate the outperformance of our approach compared to other methods. The work unlocks several directions of future research. One interesting question would be the inference problem, i.e.,\ to assess the uncertainty of the obtained estimates and the imputation. Other directions include the trade-off between (non)convex optimization and statistica efficiency. While we have provided numerical evidence for the success of nonconvex approach, the full landscape of the optimization remains open. The interplay between computational efficiency and statistical accuracy in general tensor problems warrants future research.
